### 2014-05-21 intro.tex
More work on intro

\section{Showing our work in protocol design}

@@

By ``showing our work'' in this way, clearly enough for a computer to
recreate the same design, tweaking the assumptions is a matter of
changing the inputs to a computer program and running it again. My
colleagues and I have found that this approach yields better
peformance than conventional protocols, while allowing adjacent layers
to evolve more freely.

@@

Over the last three years, my colleagues and I built a series of
systems that demonstrate the benefits of automated, objective-driven
protocol design.

\subsection{Sprout}

\textbf{Sprout} (Chapter~\ref{chap:sprout}) is a transport protocol
designed to carry high-throughput interactive traffic, such as a
videoconference, over a cellular network. Sprout includes an explicit
model of the dynamics of cellular networks and makes predictions about
future performance by inferring, with uncertainty, the current state
of the network and evolving the model forward. Its control
strategy---how much data to send at a given moment---is a function of
those predictions and of an explicit objective: maximize throughput,
but bound in-network delays to be at most 100~ms with high
probability.

@@

\subsection{Remy}

\textbf{Remy} (Chapter~\ref{chap:remy}) generalizes Sprout to address
the classical problem of \emph{multi-agent} congestion control, where
independent users contend for the same limited network resource. Remy
is a protocol-design tool that takes, as input, a set of assumptions
about the uncertain network and workload, and an objective to pursue
on behalf of the application. Remy's computer-generated algorithms can
achieve higher performance and greater fairness than some
sophisticated human-designed schemes, including ones that put
intelligence inside the network.

On a simulated 15~Mbps fixed-rate link with eight senders contending and
an RTT of 150~ms, a computer-generated congestion-control algorithm
achieved the following improvements in median throughput and
reductions in median queueing delay over these existing protocols:

@@

Once we had a computerized protocol designer, we explored using it as
a tool to gain understanding about the problem of congestion control.
How easy is it to ``learn'' a network protocol to achieve desired
goals, given a necessarily imperfect model of the networks where it
ultimately will be deployed? Is there a tradeoff between the
performance of an algorithm now---if optimized for a very specific
network---versus the protocol's ability to perform adequately as the
network evolves? What is the cost of compatibility with existing TCP
congestion-control mechanisms?

Our research into these questions is reported in
Chapter~\ref{chap:learnability}. We found weak evidence of a tradeoff
between the breadth of the operating range of a computer-generated
protocol and its performance, but also that a single
computer-generated protocol was able to outperform existing schemes
over a thousand-fold range of link rates.

We also measured the benefits and costs of adding ``TCP-awareness'' to
a computer-generated congestion-control protocol protocol. Including
the possibility of TCP cross-traffic in the network model improved
performance when TCP cross traffic was actually present---but hurt
performance when TCP cross traffic was absent, compared with a
``TCP-naive'' protocol designed only to interoperate with itself.

\begin{comment}
\subsection{Mosh}

\textbf{Mosh} (mobile shell)~\cite{mosh} is a replacement for SSH
(secure shell) in which the client and server each explicitly model
the dynamic contents of the text terminal. Unlike SSH, which transmits all
output from an application through a reliable TCP connection and
doesn't attempt to interpret the stream, Mosh's modeling allows it
to choose which data to send to the client when, in order to most
efficiently satisfy an explicit objective: update the client screen to
the current contents of the text terminal as efficiently as possible,
even across intermittent, roaming connectivity.

By having an explicit model of the user interface at both sides of the
connection, the Mosh client can predict the application's behavior in
response to user keystrokes and (if confident in its predictions)
speculatively display the result---generally an echoed keystroke or
cursor motion---before it has been confirmed by the server. In a
trace-based evaluation over 40 hours of real-world usage, Mosh was
able to successfully predict and immediately display the result of
\textbf{70\%} of user keystrokes.

Mosh is primarily a remote terminal application and is not further
discussed in this dissertation, but its design ideas---maintain an
explicit model at runtime, and pursue a stated objective on behalf of
the user---formed the basis of my subsequent work on
transport-protocol design.
\end{comment}

>>>>

%\section{Showing our work in protocol design}

@@

By ``showing our work'' clearly enough for a computer to create the
design, tweaking a protocol's assumptions becomes a matter of changing
the inputs to a computer program and running it again. My colleagues
and I have found that this approach yields better peformance than
conventional protocols, while allowing adjacent layers to evolve more
freely. It's not clear when computer-generated protocols will be
practically deployed on the broad Internet, but what computers can
teach us about protocol design can be applied to help humans create
better protocols now.

@@

My colleagues and I have built two systems that explore the benefits
of automated, objective-driven congestion-control protocol design.

\textbf{Sprout} (Chapter~\ref{chap:sprout}) is a transport protocol
designed to carry high-throughput interactive traffic, such as a
videoconference, over a cellular network. Sprout includes an explicit
model of the dynamics of cellular networks and makes predictions about
future performance by inferring a distribution over the current state
of the network and evolving the model forward. Its control
strategy---how much data to send at a given moment---is a function of
those predictions and of an explicit objective: maximize throughput,
but cap in-network delays to be at most 100~ms with high
probability.

@@

\textbf{Remy} (Chapter~\ref{chap:remy}) generalizes Sprout to address
the classical problem of \emph{multi-agent} congestion control, where
independent users contend for the same limited network resource. Remy
is a protocol-design tool that takes, as input, a set of assumptions
about the uncertain network and workload, and an objective to pursue
on behalf of the application.

On a simulated 15~Mbps link with eight senders contending and an RTT
of 150~ms, a computer-generated congestion-control algorithm achieved
the following improvements in median throughput and reductions in
median queueing delay over these existing protocols:

@@

\section{What computers can teach us about congestion control}

We can learn about transport-protocol design by observing what makes
automated protocol-design tools successful. By looking at \emph{what
  information} turned out to be valuable to Sprout and Remy, and
\emph{what behaviors} characterize the successful computer-generated
algorithms, we have been able to develop an intuition as to what
successful new designs might look like and why Sprout and Remy are
able to outperform current protocols.

\subsection{Algorithms should collect rate estimates as signals}

Today's congestion-control protocols generally pay attention to a
small number of signals derived from packets arriving at the receiver,
or acknowledgments arriving back at the sender. Classical TCP schemes,
such as NewReno~\cite{newreno} and Cubic~\cite{cubic}, attempt to
detect packet loss as a proxy for in-network buffer
overflow. Delay-based schemes such as Vegas~\cite{vegas} detect
increasing round-trip delay, as a proxy for growing standing queues
inside the network.

Researchers and practitioners have proposed that the network should
provide additional information helpful for inferring the presence of
congestion~\cite{ecn,xcp,rcp}, but these schemes have not been widely
deployed.

Sprout and Remy have found that more useful information is available
from the data already provided by today's networks. In particular,
estimates of the rates of packet transmission and reception proved to
be crucial to helping these systems outperform existing algorithms.

\begin{figure}
\caption{Sprout measures the \textbf{rate} of packet arrivals at the
  receiver, forecasts the time required to drain the current queue,
  and tries to steer this quantity not to exceed a tenth of a
  second. Based on this technique, Sprout achieves less end-to-end
  delay than the other algorithms, while achieving almost as much
  throughput as TCP Cubic. From Figure~\ref{f:allgraphs} (Verizon LTE
  Downlink).}

\def\svgwidth{\textwidth}\footnotesize\import{dotgraphs3/}{VerizonLTE-Downlink-minus1.pdf_tex}
\end{figure}
